Structured Versus Decorated Cospans

### 2020-11-17 cospans.tex

Before formally defining `pseudo double category', it is helpful to have the following picture in mind. A pseudo double category has 2-morphisms shaped like:

\[
\begin{tikzpicture}[scale=1]
\node (D) at (-4,0.5) {$A$};
\node (E) at (-2,0.5) {$B$};
\node (F) at (-4,-1) {$C$};
\node (A) at (-2,-1) {$D$};
\node (B) at (-3,-0.25) {$\Downarrow a$};
\path[->,font=\scriptsize,>=angle 90]
(D) edge node [above]{$M$}(E)
(E) edge node [right]{$g$}(A)
(D) edge node [left]{$f$}(F)
(F) edge node [above]{$N$} (A);
\end{tikzpicture}
\]

We call $A, B, C$ and $D$ \textbf{objects} or \textbf{0-cells}, $f$ and $g$ \textbf{vertical 1-morphisms}, $M$ and $N$ \textbf{horizontal 1-cells} and $a$ a \textbf{2-morphism}. Note that a vertical 1-morphism is a morphism between 0-cells and a 2-morphism is a morphism between horizontal 1-cells. We will denote both kinds of morphisms and horizontal 1-cells as a single arrow, namely `$\to$'. We follow the notation of Shulman \cite{Shulman2008} with the following definitions.

\begin{defn}\label{defn:double_category}
A \textbf{pseudo double category} $\lD$, or $\textbf{double category}$ for short, consists of a category of objects $\mathbf{D_{0}}$ and a category of arrows $\mathbf{D_1}$ with the following functors
\begin{center}
$U\maps \mathbf{D_{0}} \to \mathbf{D_1}$\\
$S,T \maps \mathbf{D_1} \rightrightarrows \mathbf{D_{0}}$\\
$\odot \maps \mathbf{D_1} \times_{\mathbf{D_{0}}} \mathbf{D_1} \to \mathbf{D_1}$ (where the pullback is taken over $\mathbf{D_1} \xrightarrow[]{T} \mathbf{D_{0}} \xleftarrow[]{S} \mathbf{D_1}$) \\
\end{center}
 such that \\
\begin{center}
$S(U_{A})=A=T(U_{A})$\\
$S(M \odot N)=SN$\\
$T(M \odot N)=TM$\\
\end{center}
equipped with natural isomorphisms
\begin{center}

>>>>

Before formally defining a `double category', it is helpful to have the following picture in mind. A double category has 2-morphisms shaped like:
\[
\begin{tikzpicture}[scale=1]
\node (D) at (-4,0.5) {$A$};
\node (E) at (-2,0.5) {$B$};
\node (F) at (-4,-1) {$C$};
\node (A) at (-2,-1) {$D$};
\node (B) at (-3,-0.25) {$\Downarrow a$};
\path[->,font=\scriptsize,>=angle 90]
(D) edge node [above]{$M$}(E)
(E) edge node [right]{$g$}(A)
(D) edge node [left]{$f$}(F)
(F) edge node [above]{$N$} (A);
\end{tikzpicture}
\]
We call $A, B, C$ and $D$ \textbf{objects}, $f$ and $g$ \textbf{vertical 1-morphisms}, $M$ and $N$ \textbf{horizontal 1-cells} and $a$ a \textbf{2-morphism}. Note that a vertical 1-morphism is a morphism between objects and a 2-morphism is a morphism between horizontal 1-cells. %We will denote both kinds of morphisms and horizontal 1-cells as a single arrow, namely `$\to$'. 
We follow the notation of Shulman \cite{Shulman2008} with the following definitions.

\begin{defn}\label{defn:double_category}
A \textbf{pseudo double category} $\lD$, or $\textbf{double category}$ for short, consists of a category of objects $\lD_0$ and a category of arrows $\lD_1$ with the following functors
\begin{center}
$U\maps \lD_0 \to \lD_1$\\
$S,T \maps \lD_1 \rightrightarrows \lD_0$\\
$\odot \maps \lD_1 \times_{\lD_0} \lD_1 \to \lD_1$ (where the pullback is taken over $\lD_1 \xrightarrow[]{T} \lD_0 \xleftarrow[]{S} \lD_1$) \\
\end{center}
 such that \\
\begin{center}
$S(U_{A})=A=T(U_{A})$\\
$S(M \odot N)=SN$\\
$T(M \odot N)=TM$\\
\end{center}
equipped with natural isomorphisms
\begin{center}
